---
title: "Stat450_project_questions"
author: "Noah Gallagher"
date: "4/27/2021"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
vegas <- read.csv('vegas2.csv',sep = ',')
library(tidyverse)
```
## Question 1 : Is there a relationship between a Hotel’s Ratings and Stars?

 a) First let us conduct ANOVA tests to analyze each variable Score and Hotel.stars
```{r}
attach(vegas)
grp = factor(Hotel.name)

# AOV test for Score
summary(aov(Score ~ grp))

# AOV test for Hotel.stars
summary(aov(Hotel.stars ~ grp))

# F stat = 2.10
g=21
p=2
N=504
alpha=0.05
qf(1-(alpha)/p, N-g, g-1)
```
-For Hotel Score: We Reject Ho meaning, all hotel ratings differ significantly among the Hotels and there is no correlation between them.
-For Hotel Stars: We Reject Ho, meaning all hotel Stars differ significantly among the Hotels.

 b) Now, let us determine the correlation between Hotel Reviews and Hotel Stars.
```{r}
# Removing NA's
vegas2<-na.omit(vegas)

attach(vegas2)

# Correlation
cor(Score, Hotel.stars)

# Fitting a line to Score and Stars
ggplot(vegas2, aes(x = Score, y = Hotel.stars))+
  geom_point(shape = 1) +
  geom_smooth(method = lm) +
  geom_jitter(shape = 1)+
  labs(y='Hotel Stars')

# Visualize correlation btw Score and Stars
subset <- vegas2%>%
  select(Score, Hotel.stars)

pairs(subset, lower.panel = NULL)
```
The correlation is 0.19, which is a weak positive correlation.

 c) Let us check proportion of points within the prediction ellipse
```{r}
#Var-cov Matrix
S = cov(subset)

# Find column means
xbar = colMeans(subset)

# Mahalanobis Test
D2 = mahalanobis(subset, xbar, S)

# Variable count
p = ncol(subset)

# Proportion
sum(D2 <= qchisq(.95, p))/nrow(subset)
```
Thus, The proportion of points within the prediction ellipse is 96.23%


 d) MANOVA TESTNG: Can we determine if there is a significant difference between the stars a hotel has and their reviews?
  (are the average of scores and average of stars are the same among all hotel names)
  Testing: Ho: Mu-Score = Mu-Stars vs. Ha: Mu-Score != Mu-Stars
```{r}
single_hotels <- vegas%>%
  select(Hotel.name, Score, Hotel.stars)%>%
  group_by(Hotel.name)%>%
  summarize(
    avg_score = mean(Score),
  )

hotel_stars <- vegas%>%
  select(Hotel.name, Hotel.stars)%>%
  group_by(Hotel.name)%>%
  summarise(
    Stars = mean(Hotel.stars)
  )

new_vegas <- left_join(single_hotels, hotel_stars)
new_vegas

#table(vegas$Hotel.name)
#attach(vegas)
Y = cbind(new_vegas$avg_score, new_vegas$Stars)
grp = as.factor(new_vegas$Hotel.name)
fit = manova(Y ~ grp)
summary(fit, test = 'Wilks')

# you can try grouping the hotel names into hotel types to reduce g.  

star_level <- new_vegas%>%
  mutate(
    
  )

high_star <- vegas%>%
  filter(Hotel.stars == 5)%>%
  mutate(
    level = 'high'
  )
high_star

med_stars <- vegas%>%
  filter(Hotel.stars == 4)%>%
  mutate(
    level = 'med'
  )

low_stars <- vegas%>%
  filter(Hotel.stars == 3)%>%
  mutate(
    level = 'low'
  )
low_stars
stars<- full_join(high_star, med_stars)
stars2<- full_join(stars, low_stars)
stars2


# Y = cbind(stars2$Score, stars2$Hotel.stars)
# grp = as.factor(stars2$level)
# fit = manova(Y ~ grp)
# summary(fit, test = 'Wilks')


```

Getting Score means for each hotel
```{r}
#attach(vegas)

score_means<-vegas%>%
  group_by(Hotel.name)%>%
  summarize(
    avg_rating = mean(Score)
  )%>%
  arrange(desc(avg_rating))
score_means

cat("The avergae score a hotel is given is: ", mean(score_means$avg_rating), "\nThe variance across all hotel's Scores is: ", var(score_means$avg_rating))

#plotting scores
ggplot(vegas, aes(x=Score)) +
            geom_histogram(bins=5) +
            labs(x='Scores', y='count')
```


Conclusion:

-all hotel ratings differ significantly among the Hotels and there is no correlation between them.
-all hotel Stars differ significantly among the Hotels.
-The correlation is 0.19, which is a weak positive correlation.
-The proportion of points within the prediction ellipse is 96.23%


## Question 2 - Can we determine if there is a significant difference between tourist reviews and USA resident reviews for hotels?

 a) Plotting difference of Means
```{r}
# Tourist Dataset
tourist_reviews<-vegas%>%
  filter(User.country != 'USA')%>%
  group_by(Score)%>%
  mutate(
    count = n()
  )

a = ggplot(tourist_reviews, aes(x=Score)) +
            geom_histogram(bins=5)+
            geom_vline(xintercept = c(2, 5), color = "blue", size = 1)+
            labs(x='Scores', y='count', title= "Tourist Reviews")

mean(tourist_reviews$Score)

# USA dataset
usa_reviews<-vegas%>%
  filter(User.country == 'USA')%>%
  group_by(Score)%>%
  mutate(
    count = n()
  )

b = ggplot(usa_reviews, aes(x=Score)) +
            geom_histogram(bins=5) +
            geom_vline(xintercept = c(2, 5), color = "blue", size = 1)+
            labs(x='Scores', y='count', title = "USA Reviews")

mean(usa_reviews$Score)

library(gridExtra)
grid.arrange(a, b, nrow=1, ncol=2)

# Get the Confidence Intervals
quantile(tourist_reviews$Score, c(0.025, 0.975))
quantile(usa_reviews$Score, c(0.025, 0.975))



# Scatter plot
ggplot(tourist_reviews, aes(x=Score, y=count)) +
            geom_point()+
            labs(x='Scores', y='count', title = "Tourist Reviews")

ggplot(tourist_reviews, aes(x=Score, y=count)) +
            geom_point()+
            labs(x='Scores', y='count', title = "USA Reviews")


# Check Correlation
set.seed(9)
tourist_reviews2 <-vegas%>%
  filter(User.country != 'USA')%>%
  select(Score)%>%
  na.omit(tourist_reviews2)%>%
  mutate(
    type = 'tourist'
  )%>%
  sample_n(200)
tourist_reviews2

usa_reviews2 <-vegas%>%
  filter(User.country == 'USA')%>%
  select(Score)%>%
  na.omit(usa_reviews2)%>%
  mutate(
    type = 'us'
  )%>%
  sample_n(200)
usa_reviews2

scores_joined <- full_join(tourist_reviews2, usa_reviews2)
scores_joined

# Correlation between both scores
#pairs(my_scores)

# cor(my_scores$tourist,my_scores$usa)

# - analyze variable Score independently based on type of tourist


# ANOVA TEST 

# Grouping by tourist type
grp = factor(scores_joined$type)

#Conducting ANOVA Test based on Review Score
summary(aov(scores_joined$Score ~ grp))

# get F stat
g=2
p=1
N=400
alpha=0.05
qf(1-alpha, N-g, g-1)
```
Tourist Mean: 4.0976
Resident Mean: 4.1567

We do NOT reject HO, all hotel scores DO NOT differ significantly among the tourist type.


## Question 3: Are the average of scores and average of stars are the same among each Period of Stay?

Lets first find the most frequent period of stay.
```{r}
period_stay <-vegas%>%
  select(Period.of.stay)%>%
  group_by(Period.of.stay)%>%
  summarize(
    num_guests = n()
  )%>%
  arrange(desc(num_guests))
period_stay
```

Now lets use MANOVA to answer: Are the average of scores and average of stars are the same among each Period of Stay?
```{r}
#attach(vegas)
Y = cbind(vegas$Score, vegas$Hotel.stars)
grp = factor(vegas$Period.of.stay)
fit = manova(Y ~ grp)
summary(fit, test = 'Wilks')

# a = 500, b = 2, c = 2
a <- N-g-((p-g+2)/2)
b <- sqrt((p^2*(g-1)^2-4) / (p^2+(g-1)^2-5))
c<- (p*(g-1)-2)/(2)

# F stat = 13.9
g=4
p=2
N=504
alpha=0.05
qf(1-(alpha), p*(g-1), (a*b)-c)

```

We do not reject Ho, average of scores and average of stars are the same among each Period of Stay.


Count of travelers
```{r}
popular_traveler <- vegas%>%
  group_by(Traveler.type)%>%
  summarize(count = n())%>%
  arrange(desc(count))

popular_traveler
```
- Most popular Traveler type is Couples.

What period of stay is best for travelers?
```{r}
#get period per traveler type count
ppt <- vegas%>%
  group_by(Traveler.type, Period.of.stay)%>%
  summarize(count = n())%>%
  rename(
    traveler = Traveler.type,
    stay = Period.of.stay
  )%>%
  arrange(traveler)
ppt

library(sqldf)
sqldf("SELECT distinct traveler, stay, max(count) as num_guests 
       FROM ppt 
       GROUP BY traveler 
       ORDER BY num_guests desc")



```
The best time for couples is: Sep-Nov
The best time for families is: Jun-Aug
The best time for Friends is: Mar-May
The best time for Business is: Dec-Feb
The best time for individuals is: Jun-Aug

## Question 4: Amenities

First lets determine What amenities are normally included for all hotels?
```{r}
# Lets see what hotels have certain amenities
normal_amenities <- vegas%>%
  group_by(Hotel.name)%>%
  summarize( 
    pools = sum(Pool == 'YES')/24, #we divide by 24 since there are 24 entries for each hotel
    gyms = sum(Gym == 'YES')/24,
    tennis = sum(Tennis.court == 'YES')/24,
    spas = sum(Spa == 'YES')/24,
    casinos = sum(Casino == 'YES')/24,
    wifi = sum(Free.internet == 'YES')/24
  )
#normal_amenities

# Get the number of hotels in dataset (21 hotels)
count_hotels <- vegas%>%
  summarize(num_hotel = length(unique(Hotel.name)))
#count_hotels

# Lets get the proportions of amenities
hotel_props<- normal_amenities%>%
  summarize(
    prop_pools = sum(pools) / 21,
    prop_gym = sum(gyms) / 21,
    prop_tennis = sum(tennis) / 21,
    prop_spas = sum(spas) / 21,
    prop_casinos = sum(casinos) / 21,
    prop_wifi = sum(wifi) / 21
  )

hotel_props
```
Thus we can conclude:
  - 95.24% of hotels have Pools
  - 95.24% of hotels have Gyms
  - 23.81% of hotels have Tennis Courts
  - 76.19% of hotels have Spas
  - 90.48% of hotels have Casinos
  - 95.24% of hotels have Free Internet
  
Let's plot these:
```{r}
# Plotting Pool Presence
pool_plot = ggplot(vegas, aes(x=Pool)) +
            geom_bar() +
            labs(x='Presence of Pool', y='count')
#pool_plot

# Plotting Gym Presence
gym_plot = ggplot(vegas, aes(x=Gym)) +
            geom_bar() +
            labs(x='Presence of Gym', y='count')
#gym_plot

# Plotting Spa Presence
spa_plot = ggplot(vegas, aes(x=Spa)) +
            geom_bar() +
            labs(x='Presence of Spa', y='count')
#spa_plot

# Plotting Tennis Court Presence
tennis_plot = ggplot(vegas, aes(x=Tennis.court)) +
            geom_bar() +
            labs(x='Presence of Tennis Court', y='count')
#tennis_plot

# Plotting Casino Presence
casino_plot = ggplot(vegas, aes(x=Casino)) +
            geom_bar() +
            labs(x='Presence of Casino', y='count')
#casino_plot

# Plotting Free Wifi
wifi_plot = ggplot(vegas, aes(x=Free.internet)) +
            geom_bar() +
            labs(x='Presence of Wifi', y='count')
#wifi_plot

# Merging
grid.arrange(pool_plot, gym_plot, spa_plot, tennis_plot, casino_plot, wifi_plot, ncol=3, nrow=2)
```

## Question 5: What correaltions can we determine?

```{r}
numeric_subset<- vegas2%>%
  select(Nr..reviews, Nr..hotel.reviews, Helpful.votes, Score, Hotel.stars, Nr..rooms, Member.years)

cor(numeric_subset)
pairs(numeric_subset, lower.panel = NULL,)
#cov(numeric_subset)
#colMeans(numeric_subset)
```
Highest: Number of reviews a user has done and the number of helpful votes they have recieved 0.79
Lowest: 
